%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Advantages of Fourier-based time methods}
\label{cha:advantages}

\chabstract{In this chapter, the advantage of using the 
harmonic balance approach to estimate the 
temporal derivative is assessed. First, an analytical
function, whose derivative is known, is chosen to
compare the accuracy of the estimation of the derivative.
using the harmonic balance operator
and two finite-difference schemes. It is shown that 
the HB operator is spectral accurate. The same
function is then used along with the linear advection toy problem
to ensure that this property is kept when iteratively solving
an equation. Finally, a function composed of two segregated frequencies
is tested and it is shown that the multi-frequential approach
is more suited to this sort of problems and gives a CPU
gain, compared to a mono-frequential HB method, that is proportional
to the segregation of the frequencies.}

\minitoc
\newpage


\section{Comparison of the harmonic balance operator and finite difference schemes}
\label{sec:hb_operator}
\input{PARTS/ADV_LIM/ADV/hb_operator}

\section{Periodic advection of a sum of sine functions}
\label{sec:sum_sine}
\input{PARTS/ADV_LIM/ADV/sum_sine}

\section{Capturing a segregated multi-frequential signal}
\label{sec:adv_multifreq}
\input{PARTS/ADV_LIM/ADV/multifreq}

\chconclu{The advantages of Fourier-based time methods
is shown in this chapter. When dealing with periodic functions
that have some regularity properties, Fourier-based time methods
(namely here the harmonic balance) are spectral accurate: they are
able to capture a signal with the smallest error by using a finite
number of harmonics. When dealing with a multi-frequential signal whose
frequencies is largely segregated, the use of the multi-frequential
version of the harmonic balance gives a tremendous gain in CPU time as
only the frequencies involved in the computation have to be set.
In opposite, the mono-frequential (and also the classical time-marching schemes)
need to discretize all the intermediate frequencies. Hence the choice of the multi-frequential
harmonic balance approach for this work.}
