%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Isolated low-speed CROR configuration}
\label{cha:dream_ls_isolated}

\chabstract{The studies performed in the previous chapters 
are finally used together to simulate the aeroelasticity
of a low-speed CROR configuration. First,
the steady results are analyzed to provide insight into the flow
physics and give confidence in the results. The prediction tool
defined in Chap.~\ref{cha:limitations_convergence}
is then used to estimate the number of harmonics required to
simulate the unsteady rigid-motion response 
of the CROR using the harmonic balance approach.
Aeroelastic simulations are then carried-out using the decoupled
approach that has been validated in the previous chapter. The OPT
algorithm develop in Chap.~\ref{cha:limitations_condition_number}
is used to ensure the stability of the computations.
Local excitation contours and the integrated damping are finally
analyzed.}


\newpage

\section{Presentation of the case}
\label{sec:dream_presentation}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/presentation}

\section{Numerical setup}
\label{sec:dream_ls_numerical}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/numerical}

\section{Steady results}
\label{sec:dream_ls_steady_results}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/steady_results}

\section{Spectral convergence of the harmonic balance computations}
\label{sec:dream_ls_spectral_convergence}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/spectral_convergence}

\section{Unsteady rigid-motion results}
\label{sec:dream_ls_rigid_results}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/rigid_results}

\section{Aeroelastic results}
\label{sec:dream_ls_ael_results}
\input{PARTS/APPLI/DREAM_ISOLATED_LS/ael_results}

\chconclu{The multi-frequential harmonic balance
approach along with the decoupled approach has
been used in this chapter to simulate the flutter
behavior of a low-speed CROR computation. It is shown
that the local excitation varies in correlation
with the inflection lines of the modes and with 
a change in aerodynamic behavior. This configuration is
shown to be cleared from flutter as the damping
is positive for all modes and all inter-blade phase angles.
To further assess the approach, a more demanding 
configuration is studied, namely a high-speed
CROR.}
