%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{11\texorpdfstring{\textsuperscript{th}}{th} standard aeroelastic configuration}
\label{cha:stcf11}

\chabstract{The harmonic balance method along with an aeroelastic 
\replaced{decoupled}{weak-coupling} approach, is applied to
the well-known 11\textsuperscript{th} standard aeroelastic configuration of
\citet{Fransson1999}. It is shown that by using only
one harmonic ($N=1$), the damping curve of both the subsonic
and the transonic operating points are superimposed with
the reference unsteady computation. The agreement
with both the experimental and the numerical data available
is good, justifying the proposed approach. Moreover, a
speed-up of seven is found compared to a classical 
time-marching scheme. This work has been published in
\begin{quote}
	\citetalias{Sicot2014}
\end{quote}}


\newpage

\section{Presentation of the case}
\label{sec:stcf11_presentation}
\input{PARTS/APPLI/STCF11/presentation}

\section{Numerical setup}
\label{sec:stcf11_numerical}
\input{PARTS/APPLI/STCF11/numerical}

\section{Subsonic case}
\label{sec:stcf11_subsonic}
\input{PARTS/APPLI/STCF11/subsonic}

\section{Transonic case}
\label{sec:stcf11_transonic}
\input{PARTS/APPLI/STCF11/transonic}

\chconclu{The use of the harmonic balance approach
to compute turbomachinery aeroelasticity has been validated
against a well documented test case. The results
are in good agreement with the experimental data and with
the numerical results obtained by using different approaches.
The harmonic balance approach
has shown that with only one harmonic, the damping curve
is retrieved compared to a classical time-marching scheme.
The speed-up obtained is seven compared to a phase-lag
approach with a classical time-marching
scheme. \added{Even though the aeroelasticity validation case is confined
to 2D, which is the case for most of the one found in the literature,} 
this give us faith to
apply the current approach to more demanding aeroelastic
computations, namely contra-rotating open rotor
aeroelasticity.}
