%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Presentation of the toy problems}
\label{cha:toy_problems}

\chabstract{In this chapter, three toy problems are
presented. They are set up to thoroughly study the properties
of the chosen harmonic balance approach defined in Sec.~\ref{sec:sm_hb}. 
They will be used all along Part~\ref{part2}.
The first
toy problem concerns the resolution of the linear advection
equation with imposed periodic perturbations. An
analytic solution is known, allowing to accurately assess
the harmonic balance method from a theoretical point of view.
The second toy problem is a channel flow with an
oscillating back-pressure. It is solved using the 2D Navier--Stokes
equations. This allows to highlight the properties of the harmonic
balance method within a non-linear equation framework. 
Finally, the third toy problem is a model problem of a turbomachinery configuration.
A Gaussian function representative of a wake is injected and convected through
the domain using the 3D Euler equations. This test case
will be used to assess the 
convergence of Fourier-based time methods 
for turbomachinery wake passing problems.
Furthermore, 
the solver used for these last two cases is the \emph{elsA}
code which will be used for the CROR applications.}

\minitoc
\newpage


\section{Periodic linear advection}
\label{sec:toy_convection}
\input{PARTS/ADV_LIM/TOY_PB/convection}

\section{Channel flow with oscillating back pressure}
\label{sec:toy_channel}
\input{PARTS/ADV_LIM/TOY_PB/channel}

\section{Model turbomachinery configuration}
\label{sec:model_tbm}
\input{PARTS/ADV_LIM/TOY_PB/model_tbm}

\chconclu{\mytodo{TODO}}
