%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Isolated high-speed CROR configuration}
\label{cha:dream_hs_isolated}

\chabstract{To further assess the proposed
multi-frequential harmonic balance method along
with a decoupled approach, a more demanding
case is studied, namely a high-speed CROR configuration.
The number of harmonics required to compute such a
configuration is shown to be higher than the 
low-speed configuration. The unsteady rigid-motion
computations are analyzed and aeroelastic
simulations are then carried out, showing that 
the proposed approach is robust enough to 
assess such configurations.}

\newpage

\section{Presentation of the case}
\label{sec:dream_hs_presentation}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/presentation}

\section{Numerical setup}
\label{sec:dream_hs_numerical}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/numerical}

\section{Steady results}
\label{sec:dream_hs_steady_results}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/steady_results}

\section{\emph{A priori} estimate of the required number of harmonics}
\label{sec:dream_hs_spectral_convergence}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/spectral_convergence}

\section{Unsteady rigid-motion results}
\label{sec:dream_hs_rigid_results}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/rigid_results}

\section{Aeroelastic results}
\label{sec:dream_hs_ael_results}
\input{PARTS/APPLI/DREAM_ISOLATED_HS/ael_results}

\chconclu{The multi-frequential harmonic balance
method along with a decoupled approach has been
assessed on a high-speed CROR configuration.
The steady computations reveal highly non-linear 
features, namely shocks. The prediction tool
defined in Chap.~\ref{cha:limitations_convergence}
is then used to estimate the number 
of harmonics required to simulate the configuration.
Seven harmonics are needed for the high-speed configuration
whereas only four were required for the low-speed one
for the same accuracy.
With no need for trial and error simulations 
to select the number of harmonics, results
are shown to be consistent, namely tip vortices and wake interactions
are well captured. 
Aeroelastic simulations are then undertaken.
The results are scrutinized with focus on the integrated damping
and the local excitation of the blades.
This configuration is shown to be flutter free, 
similarly to the low-speed one.
The proposed methodology is thus demonstrated to be able to
tackle demanding industrial configurations.}
