%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Contra-rotating open rotors}
\label{cha:cror}

\chabstract{In this chapter, we first recall the thrust
and propulsive efficiency equations. Using them,
the propeller engines are shown to be good candidates
for efficient alternative engines, mainly due to a high bypass ratio.
Geometry, general principles, similarity coefficients
and main physical phenomena of such engines are 
described. In addition to that, it is shown that even if efficient,
propeller engines suffer from a residual swirl motion.
To tackle this problem, the contra-rotating open rotor
technology is presented along with its main source of unsteadiness.
The challenges associated to this engine are finally detailed
and we show that aeroelasticity needs to be accounted for.}


\newpage

\section{Generalities of propulsion}
\label{sec:cror_intro}
\input{PARTS/REVIEW/CROR/intro}

\section{Propellers}
\label{sec:cror_propeller}
\input{PARTS/REVIEW/CROR/propeller}

\section{Contra-rotating open rotors}
\label{sec:cror_cror}
\input{PARTS/REVIEW/CROR/cror}

\section{Unsteadinesses}
\label{sec:cror_unsteady}
\input{PARTS/REVIEW/CROR/unsteady}

\section{Challenges}
\label{sec:cror_challenges}
\input{PARTS/REVIEW/CROR/challenges}


\chconclu{The concept of contra-rotating open rotor has
been presented along with the basic flow phenomena that develop in it.
The unsteady phenomena that develops through this
engine are mostly correlated with the blade passing frequency, except for
the installation effects and the non-uniform inflow. The challenges
associated with this type of engine are recalled and it is highlighted 
that aeroelasticity of such systems remain to be accounted for. This is
why the present work will focus on aeroelasticity, which is
introduced in the following chapter.}
