%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Fourier-based methods}
\label{cha:spectral_methods}

\chabstract{The four main
Fourier-based methods are presented in this chapter: the Linearized 
Unsteady Reynolds-averaged Navier-Stokes (LUR), 
the Non-Linear Harmonic (NLH), 
the Non-Linear Frequency Domain (NLFD) 
and the Harmonic Balance (HB) methods. The LUR
method comes from a linearization of the governing equations
while the three others are built to take into account for the 
non-linearities. The NLH, NLFD and HB methods
rely on a decomposition of the variable of interest
in Fourier series. By truncating these at order $N$,
$2N+1$ steady equations coupled by
a source term are obtained. 
Emphasis is put on the development
of the multi-frequential formulation and its
mathematical background to allow multi-frequential applications.
This is the case, for instance, of a pylon-rotor-rotor configuration
(namely an installed CROR) or a CROR with a moving blade, which is
the purpose of the current thesis.
The applicability of 
these methods is demonstrated in the literature
through analytical test cases, $2$D/$3$D academic 
turbomachine configurations,
industrial subsonic/transonic multi-stage applications, 
aeroelastic configurations and even unsteady
optimization problems. The cost of the methods
is almost $2N+1$ times the cost of a steady
computation with $N$ being the number of computed harmonics.}

\minitoc
\newpage

% ================
% = INTRODUCTION =
% ================
\section{Introduction}
\label{sec:sm_intro}
\input{PARTS/REVIEW/hb/intro}

\section{The \underline{L}inearized 
\underline{U}nsteady \underline{R}eynolds-averaged Navier-Stokes method (LUR)}
\label{sub:sm_lur}
\input{PARTS/REVIEW/HB/lur}

\section{The \underline{N}on-\underline{L}inear 
\underline{H}armonic method (NLH)}
\label{sub:sm_nlh}
\input{PARTS/REVIEW/HB/nlh}

\section{The \underline{N}on-\underline{L}inear 
\underline{F}requency \underline{D}omain method (NLFD)}
\label{sub:sm_nlfd}
\input{PARTS/REVIEW/HB/nlfd}

% ====================
% = HARMONIC BALANCE =
% ====================
\section{The \underline{H}armonic \underline{B}alance method (HB)}
\label{sec:sm_hb}
\input{PARTS/REVIEW/HB/hb}

% ===================================
% = Periodic flows in turbomachines =
% ===================================
\section{Periodic flows in turbomachines}
\label{sec:sm_hudson}
\input{PARTS/REVIEW/HB/hudson}

\chconclu{Four Fourier-based methods have been presented in this chapter.
The main mathematical development have been demonstrated and 
the hypothesis/weakness of each method has been highlighted.
In particular, the presented harmonic balance method can
treat both mono and multi-frequential applications. One
unsteady equation is transformed into a subset of $2N+1$
steady-state equations coupled by a source term. This last
is analytical in the mono-frequential formulation and
of matrix form in the multi-frequential framework.
The large literature available on these methods shows that
they are ready for industrial, numerically hard
unsteady applications. Moreover, the method is able to
efficiently compute aeroelastic phenomenon on 
contra-rotating configurations explaining its
use in the present thesis.}
