%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Introduction to aeroelasticity}
\label{cha:ael}

\chabstract{In this chapter, the basic elements
to understand aeroelasticity in turbomachinery and by extension
in contra-rotating open rotors are detailed. Firstly, the
definition and the basic equations governing dynamic
aeroelasticity are presented. The two main aeroelastic
phenomena that develop in turbomachinery,
namely forced response and flutter, are then presented.
The latter is investigated in this work and the computational
approach retained to simulate it, namely the \replaced{decoupled}{weak-coupling} approach,
is presented. The variables that are used to quantify the 
flutter boundary are finally presented.}


\newpage

\section{What is aeroelasticity}
\label{sec:what_is_ael}
\input{PARTS/REVIEW/AEL/what_is_ael}

\section{Main aeroelastic phenomena in turbomachinery}
\label{sec:ael_phenomena}
\input{PARTS/REVIEW/AEL/ael_phenomena}

\section{\texorpdfstring{\underline{C}}{C}omputational 
\texorpdfstring{\underline{A}}{A}ero\texorpdfstring{\underline{E}}{E}lasticity (CAE)}
\label{sec:cae}
\input{PARTS/REVIEW/AEL/cae}

\chconclu{Contra-rotating open rotors are made of
thin rotor blades that see unsteadinesses coming from
the opposite rotor. The study of the coupling between 
the structural 
displacement and the unsteady flow field is called dynamic
aeroelasticity. Two approaches exist to simulate such a phenomenon:
the strong-coupling and the \replaced{decoupled one}{weak-coupling}. Even though
strong-coupling remains more accurate, the cost
of such an approach prevents its use in the industry.
Therefore, the decoupled approach is the method 
that will be used in this work to
assess the flutter of contra-rotating open rotors.
It has been described in this chapter. The 
formula to compute the damping within the \replaced{decoupled}{weak-coupling}
approach has been given. In the next chapter, Fourier-based
time methods will be presented as they are good candidates
to efficiently simulate the aeroelasticity of contra-rotating
open rotors within the \replaced{decoupled}{weak-coupling}
approach.}
