%!TEX root = ../../adrien_gomar_phd.tex
\chapter{Spectral methods}
\label{cha:spectral_methods}

\chabstract{The four main
spectral methods are presented in this chapter: the Linearized 
Unsteady Reynolds-averaged Navier-Stokes (LUR), 
the NonLinear Harmonic (NLH), 
the NonLinear Frequency Domain (NLFD) 
and the Harmonic Balance (HB) methods. The LUR
method comes from a linearization of the governing equations
while the three others are built to take into account for the 
non-linearities. The NLH, NLFD and HB methods
rely on a decomposition of the variable of interest
in Fourier series. By truncating these at order $N$,
$2N+1$ steady equations coupled by
a source term are obtained. 
Emphasis is put on the development
of multi-frequential formulation and its
mathematical background to allow multi-stage
configurations.
The applicability of 
these methods is demonstrated in the literature
through analytical test cases, $2$D/$3$D academic 
turbomachine configurations,
industrial subsonic/transonic multi-stage applications, 
aero-elastic configurations and even unsteady
optimization problems. The cost of the methods
is almost $2N+1$ times the cost of a steady
computation with $N$ being the number of computed harmonics.}

\minitoc

% ================
% = INTRODUCTION =
% ================
\section{Introduction}
\label{sec:sm_intro}
\input{CHAPTERS/hb/intro}

\input{CHAPTERS/hb/state_of_the_art}

\section{The \underline{L}inearized 
\underline{U}nsteady \underline{R}eynolds-averaged Navier-Stokes method (LUR)}
\label{sub:sm_lur}
\input{CHAPTERS/HB/lur}

\section{The \underline{N}on\underline{L}inear 
\underline{H}armonic method (NLH)}
\label{sub:sm_nlh}
\input{CHAPTERS/HB/nlh}

\section{The \underline{N}on\underline{L}inear 
\underline{F}requency \underline{D}omain method (NLFD)}
\label{sub:sm_nlfd}
\input{CHAPTERS/HB/nlfd}

% ====================
% = HARMONIC BALANCE =
% ====================
\section{The \underline{H}armonic \underline{B}alance method (HB)}
\label{sec:sm_hb}
\input{CHAPTERS/HB/hb}

% ===================================
% = Periodic flows in turbomachines =
% ===================================
\section{Periodic flows in turbomachines}
\label{sec:sm_hudson}
\input{CHAPTERS/HB/hudson}

\chconclu{Four spectral methods have been presented in this chapter.
The main mathematical development have been demonstrated and 
the hypothesis/weakness of each method has been highlighted.
In particular, the presented harmonic balance method can
treat both mono and multi-frequential applications. One
unsteady equation is transformed into a subset of $2N+1$
steady-state equations coupled by a source term. This last
is analytical in the mono-frequential formulation and
of matrix form in the multi-frequential framework.
The large literature available on these methods shows that
they are ready for industrial, numerically hard
unsteady applications. Moreover, the method is able to
efficiently compute aero-elastic phenomenon on 
contra-rotating configurations explaining its
use in the present thesis.}
