%!TEX root = ../../../adrien_gomar_phd.tex
\chapter{Preliminary validations of the harmonic balance approach}
\label{cha:validation_hb}

\chabstract{In this chapter the mono- and
multi-frequential harmonic balance approaches
are validated. In this aim, two model problems are used.
Firstly, the harmonic balance method is applied to the linear advection equation 
supplemented with unsteady boundary conditions 
of different frequency content. Mono- and multi-frequential
unsteady signals are injected and compared to
exact solutions. The spectral accuracy of the method is verified.
Moreover, the ability of the multi-frequential
approach to capture signal composed of segregated frequencies
is underlined. Based on these results, the aeroelasticity
of contra-rotating open rotors is put into perspectives.
Secondly, the multi-frequential approach is assessed for
a channel flow problem solved using the \emph{elsA}~\cite{Cambier2013}
CFD code within the Navier--Stokes equations framework. The results
are shown to be superimposed with a classical time-marching solution.}

\newpage

\section{Linear advection of a periodic signal}
\label{sec:linear}
\input{PARTS/ADV_LIM/VALIDATION_HB/linear}

\section{Application to a non-linear system of equations}
\label{sec:non_linear}
\input{PARTS/ADV_LIM/VALIDATION_HB/non_linear}

\chconclu{The mono- and multi-frequential harmonic balance
approaches have been thoroughly validated in this chapter. 
If the provided number of harmonics is sufficient, the solution is
superimposed on the exact/reference solution for both linear
and non-linear test cases. This property is called spectral accuracy.
In addition to that, it is shown that the multi-frequential harmonic
balance approach is a good candidate for the simulation of problems
involving segregated frequencies. Actually, it is shown that this is the case
of contra-rotating open rotor aeroelasticity. In the following chapter,
we will see that an issue rises when considering segregated frequencies
with the multi-frequential harmonic balance approach. As this is likely
to be the case for contra-rotating open rotor aeroelasticity, an original
approach is proposed to alleviate this problem.}
